### 2014-08-28 D1-3.tex
Renamed file. Inserted rough TOC.

\documentclass[external]{20120615_deliverable_template_ukob}

\usepackage{xspace}
\usepackage{amsmath,amsthm,amssymb,url,color}
\theoremstyle{definition}
\newtheorem{example}{Example}
\usepackage{xcolor}
\usepackage{color, colortbl}
%\usepackage{showkeys}

\LGtitle{\LiveGovThirtyTitleFonts \textbf{D1.3 - Privacy Aware Sensor Data Storage and Miner} }
%\LGnumber{WP.NR} % e.g 6.7 no letters D or ID

\LGnumber{1.2}

\LGwp{WP1 - Reality Sensing and Mining}

\LGdissemination{PU - Public}

\LGcontractdate{Month 30, October 2014}

\LGactualdate{Month 30, October 2014}

\LGtask{T1.3, T1.4}

\LGtype{Prototype}

%\LGnature{$<$ Report $|$ Prototype $|$ Demonstrator $|$ Other $>$}

%\LGapproved % activate only if approved
\LGfinal

\LGversion{12}
%\LGstaffmonths{Resources spent on the deliverable}

%\LGdistribution{WP leaders, PMB members, European Commission}

%\LGkeywords{}

%\LGabstract{Put abstract here (separate paragraphs using
%{\tt\char92 par}.)}
\LGabstract{

% \TODO{Add abstract}

%  This deliverable describes the mobile sensor mining component.
%  Along with this document we supply the source code of the
%  components, documentation (javadoc), and pre-compiled packages for
%  direct installation and testing on mobile devices.

}

%\LGhistory{01}{2011-10-01}{First draft}{Yiannis Kompatsiaris}
%\LGhistory{02}{2011-10-02}{Modifications}{Sotiris Diplaris}
%\LGhistory{03}{2011-10-03}{Further Modifications \& Corrections}{Sotiris Diplaris}
%\LGhistory{04}{2011-10-05}{Further Modifications \& Corrections}{Sotiris Diplaris}

%------------------------ macro for change portrait to landscape
\usepackage{calc,graphicx,pdflscape,eso-pic} % needed for printing
                                             % headers and footers
\newlength\landscapewidth
\newlength\landscapeheight
\newcommand\landscapepagestyle{ % command which prepare page
                                % for landscape printing,
                                % change to landscape is
                                % achieved by pdflscape
\clearpage \thispagestyle{empty}
\setlength\landscapewidth{247mm}
\setlength\landscapeheight{161mm}
\AddToShipoutPicture*{\AtPageCenter{ % this make "layer" for
                                     % printing the header and footer
\rotatebox[origin=c]{90}{
\hspace*{-2em} % for adjusting position of
               % header and footer
\parbox{\landscapewidth}{\vskip-.57\landscapeheight
%\centerline{\WEKNOWIT \hfill \textbf{\Large \bf{IDx.x - V0} \normalsize }} %header text for Internal Deliverables
                                                                         % notation: IDx.x V{LGversion}
\centerline{\LiveGovLogo \hfill \textbf{\Large \bf{Dx.x - V0} \normalsize }} %header text for External Deliverables
                                                                        % notation: Dx.x V{LGversion}
\vspace{-5pt} \rule{\landscapewidth}{0.4pt} \\
\rule{0pt}{\landscapeheight} \\
%\rule{\landscapewidth}{0.4pt} \\
\centerline{Page \thepage\ }  % footer text
} } } } }

%--------------------------------------------------------------------------------------------------------------%
\usepackage{appendix}
\pretolerance=10000 % prevent overflow lines for $math$ elements

\begin{document}

% add a \LGaddhistory{version}{date}{reason}{revised by} for each new
% version
\begin{LGhistory}
\LGaddhistory{01}{2014-07-17}{Outline}{Heinrich Hartmann}

\end{LGhistory}

\newcommand{\LGaddauthorNoPhone}[3]{\hline  #1 &  #2 & %
   \parbox{3em}{E-mail:} \small #3 \\
}

% add a
% \LGaddauthor{Partner}{Name}{Telephone}{Fax}{Email} for each author
\begin{LGauthors}

\LGaddauthor{UKob}{Heinrich Hartmann}{+49 261 287 2759}%
{+49 261 287 100 2759}{\small hartmann@uni-koblenz.de}

\LGaddauthor{UKob}{Christoph Schaefer}{+49 261 287 2786 }%
{+49 261 287 100 2786}{\small chrisschaefer@uni-koblenz.de}

%% \LGaddauthor{UKob}{Christoph Kling}{+49 261 287 2702 }%
%% {+49 261 287 100 2702}{\small ckling@uni-koblenz.de}

%% \LGaddauthorNoPhone{MTS}{Laura Niittyl\"a}%
%% {\small Laura.Niittyla@mattersoft.fi}

\end{LGauthors}

\begin{LGExecutiveSummary}
  \vspace{10pt}

  The deliverable is accompanied with source code of the components in
  Java, API documentation (javadoc), and pre-compiled packages for
  direct installation and testing on mobile devices.

\end{LGExecutiveSummary}

% add a \LGaddabbreviation{ABBR}{Explanation} for each abbreviation
\begin{LGAbbreviations}
%\LGaddabbreviation{LG}{Live+Gov}

\LGaddabbreviation{\textbf{AIDL}}
{Android Interface Description Language}

\LGaddabbreviation{\textbf{AJAX}}
{Asynchronous JavaScript and XML}

\LGaddabbreviation{\textbf{ALC}}
{Attribute Language with Complement}

\LGaddabbreviation{\textbf{API}}
{Application Programming Interface}

\LGaddabbreviation{\textbf{GeSA}}
{Geographical Semantic Analysis}

\LGaddabbreviation{\textbf{GPS}}
{Global Positioning System}

\LGaddabbreviation{\textbf{GSM}}
{Global System for Mobile Communications}

\LGaddabbreviation{\textbf{HTML}}
{HyperText Markup Language}

\LGaddabbreviation{\textbf{HTTP}}
{Hypertext Transfer Protocol}

\LGaddabbreviation{\textbf{ID}}
{Identifier}

\LGaddabbreviation{\textbf{IEEE 802.11}}
{see WIFI, WLAN}

\LGaddabbreviation{\textbf{JSON}}
{JavaScript Object Notation}

\LGaddabbreviation{\textbf{LAN}}
{Local Area Network}

\LGaddabbreviation{\textbf{REST}}
{Representational State Transfer}

\LGaddabbreviation{\textbf{RF-ID}}
{Radio-Frequency Identification}

\LGaddabbreviation{\textbf{SDCF}}
{Sensor Data Collection Framework}

\LGaddabbreviation{\textbf{SQL}}
{Structured Query Language}

\LGaddabbreviation{\textbf{SVM}}
{Support Vector Machine}

\LGaddabbreviation{\textbf{TBox}}
{Terminological Box}

\LGaddabbreviation{\textbf{UI}}
{User Interface}

\LGaddabbreviation{\textbf{URL}}
{Uniform Resource Locator}

\LGaddabbreviation{\textbf{UUID}}
{Universal Unique Device Identifier}

\LGaddabbreviation{\textbf{WP}}
{Work Package}

\LGaddabbreviation{\textbf{WIFI}}
{Wireless Fidelity (IEEE 802.11), WLAN}

\LGaddabbreviation{\textbf{WLAN}}
{Wireless Local Area Network}

\LGaddabbreviation{\textbf{XML}}
{Extensible Markup Language}

\LGaddabbreviation{~\\}
{~~~}

\end{LGAbbreviations}

% add a \LGaddterm{Term}{Explanation} for each entry of the glossary
%\begin{LGGlossary}
%\LGaddterm{Term}{Definition text here}
%\LGaddterm{Term2}{Definition text here}
%\end{LGGlossary}

\setcounter{tocdepth}{1}

% add this for a table of contents
\LGTOC

\newpage
\newcommand{\todo}[2]{\textcolor{magenta}{#1: #2}}
%\newcommand{\todo}[2]{}

\chapter{Introduction}
\label{chap:Introduction}

\clearpage

\chapter{Privacy Protection}
Intro
\section{Historical Definitions}
Aristotle
Warren and Brandeis
Fried
\section{Privacy Definition and Taxonomy}
\section{Privacy and Law}
European Convention on Human Rights - Article 8
EU Data Protection Direvtive
\section{IT Security Analysis}
Live+Gov Privacy Protection Analysis
Step 1. World Analysis
Assets: Privacy
IT-Systems and Humans
Conflict of Interests
Vulnerabilities
Step 2. Potential Analysis
Step 3. Plan Development
\section{Implementation}
* Sensor Data Privacy Control
* Rising Awareness

\chapter{Improved Sensor Data Mining Methods}
\section{Battery Awareness of Sensor Collector}
Results from Projektprakikum (Wifi/Zip)

\section{Service Line Detection (new method)}
MA thesis results from Sven Milker

\section{Issue Analysis}
Christoph Schäfer with Niko Beck

\bibliography{D1-3}

\end{document}