### 2014-09-11 figures/LG-Sensor-Privacy-Matrix.tex

\begin{figure}
\centering
\includegraphics{diagrams/png/sensor-privacy-matrix.png}

\begin{flushleft}
\scriptsize
\textbf{Legend:}
\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  The x-axis lists the 7 Types of Privacy according to Friedewald et al.
\item
  The y-axis lists all mobile sensors currently used by the Live+Gov
  procject.
\item
  The big black bullet point dentotes that there is a \textbf{direct}
  violoation of a privacy type from a sensor
\end{itemize}
\end{flushleft}

\caption{Live+Gov Sensor-Privacy Matrix}
\label{figure:Live+Gov Sensor-Privacy Matrix}
\end{figure}