### 2014-09-14 figures/ThreatTable.tex
started ThreatTable with mapping to CofI and Vulnerabilities

% =================================================
% Threat Table Macros

% Const Column Width
\newcommand{\ThreatTableColWidth}{5cm}

% Const Header Row Height
\newcommand{\ThreatTableHeaderRowHeight}{0.7cm}

% Const Content Row Height
\newcommand{\ThreatTableContentRowHeight}{1cm}

% Define Header Cell Makro
\newcommand{\ThreatTableHeaderCell}[1]{
\begin{minipage}[t][\ThreatTableHeaderRowHeight][c]{\ThreatTableColWidth}
\centering
\textbf{#1}
\end{minipage}
}

% Define Content Cell Makro
\newcommand{\ThreatTableContentCell}[1]{
\begin{minipage}[t][][c]{\ThreatTableColWidth}
\begin{flushleft}
\scriptsize
#1
\newline
\end{flushleft}
\end{minipage}
}

% Define Header Row Makro
\newcommand{\ThreatTableHeaderRow}[4]{
\ThreatTableHeaderCell{#1}
&\ThreatTableHeaderCell{#2}
&\ThreatTableHeaderCell{#3}
&\ThreatTableHeaderCell{#4}
\\ \hline
}

% Define Content Row Makro
\newcommand{\ThreatTableContentRow}[4]{
\ThreatTableContentCell{#1}
&\ThreatTableContentCell{#2}
&\ThreatTableContentCell{#3}
&\ThreatTableContentCell{#4}
\\ \hline
}

%  Define Threat Table Environment
\newenvironment{ThreatTable}
{
\begin{tabular}{|c|c|c|c|}
\hline
\ThreatTableHeaderRow
{Description}
{Conflict of Interest}
{Vulnerabilites}
{Assets (Privacy Types)}
}
{\end{tabular}}

% =================================================
% Threat Table Content

\begin{landscape}
\begin{figure}
\centering
\begin{ThreatTable}

\ThreatTableContentRow
{\textbf{(Zero Day) Exploits}
\\A criminal uses (Zero Day) Exploits to obtain access to hardware or software which stores or processes
privacy sensitive data in order to get that data.}
{Criminals want to obtain access or data for personal profit, i.e. by (re-)selling the data or  by processing it themselves. 
However, criminals may have no financial interest, they could 
\\\textbf{Criminal vs Citizen}
\\
\\ \textbf{Criminal vs Local Authorities}
\\ \textbf{Criminal vs Maintenance Staff}}
{Unsecure Hardware- or Software-Interfaces}
{\textbf{Explicit:} 1,4,6,7}

\ThreatTableContentRow
{\textbf{Man in the Middle} 
\\A criminal intercepts communication between mobile device and server or between server and application.}
{arg2}
{Unencrypted hardware or software communication}
{\textbf{Explicit:} 1,4,6,7}

\ThreatTableContentRow
{\textbf{Corrupt Employees} 
\\ Employee sells private data to advertisers}
{\textbf{Employee vs Citizen} 
\\ Employees want to make personal profit by selling citizen data. 
\\ Citizens provide data for public improvement, they don't want their data to be used for other purposes.}
{Full DB access of Employees}
{\textbf{Explicit:} 1,4,6,7}

\ThreatTableContentRow
{\textbf{Corrupt Local Authorities}}
{arg2}
{arg3}
{\textbf{Explicit:} 1,4,6,7}

\ThreatTableContentRow
{\textbf{Careless Citizen}}
{arg2}
{arg3}
{\textbf{Explicit:} 1,4,6,7}

\ThreatTableContentRow
{\textbf{Intransparent Data Mining}}
{arg2}
{arg3}
{arg4}

\end{ThreatTable}

\caption{Threat Table}
\end{figure}
\end{landscape}