### 2014-09-15 figures/IT-Systems.tex

\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{diagrams/png/it-systems.png}

\begin{flushleft}
\scriptsize
\textbf{Legend:}
\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt

\item
\textbf{(a) Citizen:} 
User of the L+G client application whose privacy is at stake.

\item
\textbf{(b) System Provider:} Provides technical infrastructure.

\item
\textbf{(c) Local Authority:} Provider of the L+G system.

\item
\textbf{(d) Mobile Device:}
Runs the L+G client application, produces and stores data sensitive to the users privacy.

\item
\textbf{(e) Data Center:} Runs the L+G services, processes and stores user data.

\item
\textbf{(f) Report Tool:} Interface to aggregated user data.

\end{itemize}
\end{flushleft}

\caption{IT Systems}
\label{figure:IT Systems}
\end{figure}