### 2014-11-05 D1-3.tex

\documentclass[external]{20120615_deliverable_template_ukob}
\usepackage{xspace}
\usepackage{amsmath,amsthm,amssymb,url,color}
\theoremstyle{definition}
\newtheorem{example}{Example}
\usepackage{xcolor}
\usepackage{color, colortbl}
\usepackage{listings}
\usepackage{subfigure}
%\usepackage{showkeys}

\LGtitle{\LiveGovThirtyTitleFonts \textbf{D1.3 - Privacy Aware Sensor Data Storage and Miner} }
%\LGnumber{WP.NR} % e.g 6.7 no letters D or ID

\LGnumber{1.2}

\LGwp{WP1 - Reality Sensing and Mining}

\LGdissemination{PU - Public}

\LGcontractdate{Month 30, October 2014}

\LGactualdate{Month 30, October 2014}

\LGtask{T1.3, T1.4}

\LGtype{Prototype}

%\LGnature{$<$ Report $|$ Prototype $|$ Demonstrator $|$ Other $>$}

%\LGapproved % activate only if approved
\LGfinal

\LGversion{12}
%\LGstaffmonths{Resources spent on the deliverable}

%\LGdistribution{WP leaders, PMB members, European Commission}

%\LGkeywords{}

%\LGabstract{Put abstract here (separate paragraphs using
%{\tt\char92 par}.)}
\LGabstract{

% \TODO{Add abstract}

%  This deliverable describes the mobile sensor mining component.
%  Along with this document we supply the source code of the
%  components, documentation (javadoc), and pre-compiled packages for
%  direct installation and testing on mobile devices.

}

%\LGhistory{01}{2011-10-01}{First draft}{Yiannis Kompatsiaris}
%\LGhistory{02}{2011-10-02}{Modifications}{Sotiris Diplaris}
%\LGhistory{03}{2011-10-03}{Further Modifications \& Corrections}{Sotiris Diplaris}
%\LGhistory{04}{2011-10-05}{Further Modifications \& Corrections}{Sotiris Diplaris}

%------------------------ macro for change portrait to landscape
\usepackage{calc,graphicx,pdflscape,eso-pic} % needed for printing
                                             % headers and footers
\newlength\landscapewidth
\newlength\landscapeheight
\newcommand\landscapepagestyle{ % command which prepare page
                                % for landscape printing,
                                % change to landscape is
                                % achieved by pdflscape
\clearpage \thispagestyle{empty}
\setlength\landscapewidth{247mm}
\setlength\landscapeheight{161mm}
\AddToShipoutPicture*{\AtPageCenter{ % this make "layer" for
                                     % printing the header and footer
\rotatebox[origin=c]{90}{
\hspace*{-2em} % for adjusting position of
               % header and footer
\parbox{\landscapewidth}{\vskip-.57\landscapeheight
%\centerline{\WEKNOWIT \hfill \textbf{\Large \bf{IDx.x - V0} \normalsize }} %header text for Internal Deliverables
                                                                         % notation: IDx.x V{LGversion}
\centerline{\LiveGovLogo \hfill \textbf{\Large \bf{Dx.x - V0} \normalsize }} %header text for External Deliverables
                                                                        % notation: Dx.x V{LGversion}
\vspace{-5pt} \rule{\landscapewidth}{0.4pt} \\
\rule{0pt}{\landscapeheight} \\
%\rule{\landscapewidth}{0.4pt} \\
\centerline{Page \thepage\ }  % footer text
} } } } }

%--------------------------------------------------------------------------------------------------------------%
\usepackage{appendix}
\pretolerance=10000 % prevent overflow lines for $math$ elements

\begin{document}

% add a \LGaddhistory{version}{date}{reason}{revised by} for each new
% version
\begin{LGhistory}
\LGaddhistory{01}{2014-07-17}{Outline}{Heinrich Hartmann}

\end{LGhistory}

\newcommand{\LGaddauthorNoPhone}[3]{\hline  #1 &  #2 & %
   \parbox{3em}{E-mail:} \small #3 \\
}

% add a
% \LGaddauthor{Partner}{Name}{Telephone}{Fax}{Email} for each author
\begin{LGauthors}

\LGaddauthor{UKob}{Heinrich Hartmann}{+49 261 287 2759}%
{+49 261 287 100 2759}{\small hartmann@uni-koblenz.de}

\LGaddauthor{UKob}{Christoph Schaefer}{+49 261 287 2786 }%
{+49 261 287 100 2786}{\small chrisschaefer@uni-koblenz.de}

\LGaddauthorNoPhone{UKob}{Maximilan Meffert}%
{\small maxmeffert@uni-koblenz.de}

\LGaddauthorNoPhone{BIZ}{Maite de Arana Agiretxe}%
{\small komunikazioa.tekniko1@biscaytik.eu}

%% \LGaddauthor{UKob}{Christoph Kling}{+49 261 287 2702 }%
%% {+49 261 287 100 2702}{\small ckling@uni-koblenz.de}

%% \LGaddauthorNoPhone{MTS}{Laura Niittyl\"a}%
%% {\small Laura.Niittyla@mattersoft.fi}

\end{LGauthors}

\begin{LGExecutiveSummary}
  \vspace{10pt}

  The deliverable is accompanied with source code of the components in
  Java, API documentation (javadoc), and pre-compiled packages for
  direct installation and testing on mobile devices.

\end{LGExecutiveSummary}

% add a \LGaddabbreviation{ABBR}{Explanation} for each abbreviation
\begin{LGAbbreviations}
%\LGaddabbreviation{LG}{Live+Gov}

\LGaddabbreviation{\textbf{AIDL}}
{Android Interface Description Language}

\LGaddabbreviation{\textbf{AJAX}}
{Asynchronous JavaScript and XML}

\LGaddabbreviation{\textbf{ALC}}
{Attribute Language with Complement}

\LGaddabbreviation{\textbf{API}}
{Application Programming Interface}

\LGaddabbreviation{\textbf{GeSA}}
{Geographical Semantic Analysis}

\LGaddabbreviation{\textbf{GPS}}
{Global Positioning System}

\LGaddabbreviation{\textbf{GSM}}
{Global System for Mobile Communications}

\LGaddabbreviation{\textbf{HTML}}
{HyperText Markup Language}

\LGaddabbreviation{\textbf{HTTP}}
{Hypertext Transfer Protocol}

\LGaddabbreviation{\textbf{ID}}
{Identifier}

\LGaddabbreviation{\textbf{IEEE 802.11}}
{see WIFI, WLAN}

\LGaddabbreviation{\textbf{JSON}}
{JavaScript Object Notation}

\LGaddabbreviation{\textbf{LAN}}
{Local Area Network}

\LGaddabbreviation{\textbf{REST}}
{Representational State Transfer}

\LGaddabbreviation{\textbf{RF-ID}}
{Radio-Frequency Identification}

\LGaddabbreviation{\textbf{SDCF}}
{Sensor Data Collection Framework}

\LGaddabbreviation{\textbf{SQL}}
{Structured Query Language}

\LGaddabbreviation{\textbf{SVM}}
{Support Vector Machine}

\LGaddabbreviation{\textbf{TBox}}
{Terminological Box}

\LGaddabbreviation{\textbf{UI}}
{User Interface}

\LGaddabbreviation{\textbf{URL}}
{Uniform Resource Locator}

\LGaddabbreviation{\textbf{UUID}}
{Universal Unique Device Identifier}

\LGaddabbreviation{\textbf{WP}}
{Work Package}

\LGaddabbreviation{\textbf{WIFI}}
{Wireless Fidelity (IEEE 802.11), WLAN}

\LGaddabbreviation{\textbf{WLAN}}
{Wireless Local Area Network}

\LGaddabbreviation{\textbf{XML}}
{Extensible Markup Language}

\LGaddabbreviation{~\\}
{~~~}

\end{LGAbbreviations}

% add a \LGaddterm{Term}{Explanation} for each entry of the glossary
%\begin{LGGlossary}
%\LGaddterm{Term}{Definition text here}
%\LGaddterm{Term2}{Definition text here}
%\end{LGGlossary}

\setcounter{tocdepth}{2}

% add this for a table of contents
\LGTOC

\newpage
\newcommand{\todo}[2]{\textcolor{magenta}{#1: #2}}
%\newcommand{\todo}[2]{}

\chapter{Introduction}\label{chap:Introduction}
In recent days the importance of privacy protection has been amplified
by the reports about the mass surveillance of ordinary citizens on a
global scale by the NSA and other intelligence agencies around the
world.

While aiming at the noble cause of enhancing eParticipation using
mobile technologies, Live+Gov systems do process a large variety data
that is potentially infringing the citizens privacy. The captured data
includes personal information like name, phone numbers and email
addresses and sensor data from GPS and accelerometer sensors. Also
with some applications it is possible to gather images and textual
input from the citizen.

While the collection of this data is necessary for providing the
advanced services that Live+Gov aims to deliver, at he same time, the
available raw data can be used to draw a very detailed picture of the
private life of the citizen. For instance can GPS location tracking be
used to reveal shopping habits (e.g. when a car seller is visited) and
associations to political groups (when a meeting is
attended). Accelerometer data can be used to infer medical conditions
like walking disabilities.

Images can contain faces of nearby persons to with whom the citizen is
associated. All this data is highly sensitive to the citizens privacy
and can be used against the citizen if it falls in the wrong hands.

The great importance of protecting the citizens privacy should be
apparent from these examples. The European Union, as well as many
other countries in the past, has set out a number of directives that
regulate the collection, processing and use of privacy sensitive
data. The most important legislation on the European Level is the Data
Protection Directive 95/46/EC [1], which is also summarized in the
European Charta of Fundamental Rights [2] of EU citizens:

\begin{quote}
Article 7. Respect for private and family life.

Everyone has the right to respect for his or her private and family
life, home and communications.

Article 8. Protection of personal data.

(1) Everyone has the right to the protection of personal data
concerning him or her.

(2) Such data must be processed fairly for specified purposes and on
the basis of the consent of the person concerned or some other
legitimate basis laid down by law.  Everyone has the right of access
to data which has been collected concerning him or her, and the right
to have it rectified.

(3) Compliance with these rules shall be subject to control by an
independent authority.  The Live+Gov fully consortium complies with
this legislation and is moreover committed to convey and enact the
ethical values inherent in these texts.
\end{quote}

Privacy as Control over Private Data

The ethical aspects of privacy have been the subject of study of many
social scientists and philosophers [3]. Many researchers do focus on
the ways in which privacy can be infringed. Thus they invest a great
amount of work in exploring threats, instead of describing why privacy
is so valuable to us. One scholar which follows an alternative aproach
in the context of digital monitoring is Charles Fried [4]. He
investiages, why we are intuitively so sensitive to violations of our
privacy. For him privacy is not asserted as an intrinsic value by
itself, he rather sated:

\begin{quote}
Privacy is not simply an absence of information about us in the minds of others;
rather it is the control we have over information about ourselves.
\end{quote}

Depending on the conversation partner, we change the degree of
information we share if it is a total stranger, a colleague, close
friend or a doctor. With doctors we share information of great
intimacy we do not share with anyone else.

Moreover, we trust those friends to not reveal information about us to
others by respecting their privacy. Trust needs the possibility of
unknown failure. If we would constantly monitor our partners, they
cannot fail unnoticed nor can they willingly share that information
with us.  So privacy, according to Fried, is the foundation of our
core relations. And thus it is valuable, because those relations are
essential to human society.

Fried’s study on the understanding of privacy provided a great
contribution to the research on the same term in philosophy and
computer science [5] and despite the fact his text was published in
1970, he already included technologies to its viewpoint (like location
monitoring) that are particularly relevant to our Context.

Live+Gov Privacy Dashboard gives citizens back the Control

In order to meet the the legal requirements and offer the citiens to
keep a maximum of contol about their personal data, Live+Gov systems
implements a number of privacy protection measure. One of the measures
that is under current development is a Privacy Dashboard (Figure
1). Using this dashboard the citizen will be able to take the
following actions: View and export all data from the citizen that is
currently stored in the Live+Gov system Selectively delete parts of
the stored data Get information about processing applied to this data
Consent or decline selectively to processing steps

View the end poducts of data mining applied to his data in a
comprehensive, graphical way These measures provide the citizen with a
maximal respect for their privacy while at the same time allowing
advanced data processing. Although, abusive exploitation of the data
is still possible, the level of transparency offered by the dashboard
builds trust between the citizen and the service provider.

\input{PrivacyProtection-2014-09-09}

\clearpage

\input{Implementations}

\chapter{Summary}\label{chap:Summary}

\bibliography{D1-3}
\end{document}