### 2014-11-07 D1-3.tex
Updated intro and summary.

\documentclass[external]{20120615_deliverable_template_ukob}
\usepackage{xspace}
\usepackage{amsmath,amsthm,amssymb,url,color}
\theoremstyle{definition}
\newtheorem{example}{Example}
\usepackage{xcolor}
\usepackage{color, colortbl}
\usepackage{listings}
\usepackage{subfigure}
%\usepackage{showkeys}

\LGtitle{\LiveGovThirtyTitleFonts \textbf{D1.3 - Privacy Aware Sensor Data Storage and Miner} }
%\LGnumber{WP.NR} % e.g 6.7 no letters D or ID

\LGnumber{1.2}

\LGwp{WP1 - Reality Sensing and Mining}

\LGdissemination{PU - Public}

\LGcontractdate{Month 30, October 2014}

\LGactualdate{Month 30, October 2014}

\LGtask{T1.3, T1.4}

\LGtype{Prototype}

%\LGnature{$<$ Report $|$ Prototype $|$ Demonstrator $|$ Other $>$}

%\LGapproved % activate only if approved
\LGfinal

\LGversion{12}
%\LGstaffmonths{Resources spent on the deliverable}

%\LGdistribution{WP leaders, PMB members, European Commission}

%\LGkeywords{}

%\LGabstract{Put abstract here (separate paragraphs using
%{\tt\char92 par}.)}
\LGabstract{

% \TODO{Add abstract}

%  This deliverable describes the mobile sensor mining component.
%  Along with this document we supply the source code of the
%  components, documentation (javadoc), and pre-compiled packages for
%  direct installation and testing on mobile devices.

}

%\LGhistory{01}{2011-10-01}{First draft}{Yiannis Kompatsiaris}
%\LGhistory{02}{2011-10-02}{Modifications}{Sotiris Diplaris}
%\LGhistory{03}{2011-10-03}{Further Modifications \& Corrections}{Sotiris Diplaris}
%\LGhistory{04}{2011-10-05}{Further Modifications \& Corrections}{Sotiris Diplaris}

%------------------------ macro for change portrait to landscape
\usepackage{calc,graphicx,pdflscape,eso-pic} % needed for printing
                                             % headers and footers
\newlength\landscapewidth
\newlength\landscapeheight
\newcommand\landscapepagestyle{ % command which prepare page
                                % for landscape printing,
                                % change to landscape is
                                % achieved by pdflscape
\clearpage \thispagestyle{empty}
\setlength\landscapewidth{247mm}
\setlength\landscapeheight{161mm}
\AddToShipoutPicture*{\AtPageCenter{ % this make "layer" for
                                     % printing the header and footer
\rotatebox[origin=c]{90}{
\hspace*{-2em} % for adjusting position of
               % header and footer
\parbox{\landscapewidth}{\vskip-.57\landscapeheight
%\centerline{\WEKNOWIT \hfill \textbf{\Large \bf{IDx.x - V0} \normalsize }} %header text for Internal Deliverables
                                                                         % notation: IDx.x V{LGversion}
\centerline{\LiveGovLogo \hfill \textbf{\Large \bf{Dx.x - V0} \normalsize }} %header text for External Deliverables
                                                                        % notation: Dx.x V{LGversion}
\vspace{-5pt} \rule{\landscapewidth}{0.4pt} \\
\rule{0pt}{\landscapeheight} \\
%\rule{\landscapewidth}{0.4pt} \\
\centerline{Page \thepage\ }  % footer text
} } } } }

%--------------------------------------------------------------------------------------------------------------%
\usepackage{appendix}
\pretolerance=10000 % prevent overflow lines for $math$ elements

\begin{document}

% add a \LGaddhistory{version}{date}{reason}{revised by} for each new
% version
\begin{LGhistory}
\LGaddhistory{01}{2014-07-17}{Outline}{Heinrich Hartmann}

\end{LGhistory}

\newcommand{\LGaddauthorNoPhone}[3]{\hline  #1 &  #2 & %
   \parbox{3em}{E-mail:} \small #3 \\
}

% add a
% \LGaddauthor{Partner}{Name}{Telephone}{Fax}{Email} for each author
\begin{LGauthors}

\LGaddauthor{UKob}{Heinrich Hartmann}{+49 261 287 2759}%
{+49 261 287 100 2759}{\small hartmann@uni-koblenz.de}

\LGaddauthor{UKob}{Christoph Schaefer}{+49 261 287 2786 }%
{+49 261 287 100 2786}{\small chrisschaefer@uni-koblenz.de}

\LGaddauthorNoPhone{UKob}{Maximilan Meffert}%
{\small maxmeffert@uni-koblenz.de}

\LGaddauthorNoPhone{BIZ}{Maite de Arana Agiretxe}%
{\small komunikazioa.tekniko1@biscaytik.eu}

%% \LGaddauthor{UKob}{Christoph Kling}{+49 261 287 2702 }%
%% {+49 261 287 100 2702}{\small ckling@uni-koblenz.de}

%% \LGaddauthorNoPhone{MTS}{Laura Niittyl\"a}%
%% {\small Laura.Niittyla@mattersoft.fi}

\end{LGauthors}

\begin{LGExecutiveSummary}
\vspace{10pt}

This deliverable presents the privacy aware, server based reality
sensing infrastructure for Live+Gov applications.  Privacy protection
is a crucial point for public acceptance of advanced data mining
techniques in eParticipation scenarios.  In order to provide such
protection we perform an extensive analysis of all threats to privacy
and implement a number of measures that address these threats in an
adequate way.

The term "privacy'' itself is highly ambiguous and there is no common
understanding what privacy really is in the literature.  In
Chapter \ref{chap:privacy} we discuss different legal and
philosophical texts and about the subject.  The main relevant
legislation is the EU Data Protection Directive (cf. Section
\ref{EUDIR}). The main ethical treatment is provided by Charles Fried
(cf. Section \ref{sec:ethics}), who defines privacy as the
\emph{control over personal information}. Furthermore, seven concrete
privacy types are introduced (cf. Section \ref{sec:privacytypes}) and
the relation to collected sensor data from mobile phones is made.

On the basis of these foundations perform a thorough analysis of all
risks for the privacy of a citizen. The main difficulty we are
presented with is the selection of sensible measures. There are a lot
of anonymization techniques in the literature and it is easy to jump
ahead and implement a few of them while leaving out massive privacy
gaps at other ends. Therefore a structured approach to this problem
was taken.

We have chosen to use the IT Security Reference Model
\cite{Grimm:ItSecRefModel} (cf. Section \ref{sec:GrimmModel}), that
was developed by our colleague from Prof. Grimm from University of
Koblenz, as a framework to conduct a systematic analysis of all risks
that are posed to the privacy of the citizens. The analysis is carried
out in Section \ref{sec:PPA}. Prof. Grimm also contributed with his
expertise and guidance to the execution of this analysis.

The analysis starts by describing an abstract version of the IT
Systems and the involved actors with their interests. We then describe
vulnerabilities of the IT Systems and conflicts between the interests
of the stakeholders. All threats that are posed to the privacy of the
citizen are caused by a conflict of interest between two actors and
exploit a vulnerability of the system.

We identity six main threats that include ``T1. Insufficent Control
Features'', ``T2. Excessive Data Minig'' and ``T4. Surveillance'' and
asses their risk of occurrence. The biggest threat to the citizens
privacy is posed by Insufficient Control Features of the system. As
soon as collected data of Citizens is stored on the servers, all
control over that data is lost if the service provider does not offer
tools for the citizen to control his data.

These risks are addressed by eight recommendations, that when
implemented provide an adequate protection of the citizens privacy.
One surprising outcome, is that the main privacy threats can be
eliminated by enhancing communication with the citizen and
establishing policies and awareness for privacy aspects inside the
company. Thus, a central point in the protection of the citizens privacy is
a ``Privacy Dashboard'' that gives the citizen a maximum of control
over his data. Using this dashboard he is able to view and export all
data from the citizen that is currently stored in the Live+Gov system,
selectively delete parts of the stored data and get information about
processing applied and view the processing results themselves.

In addition, we have implemented four different anonymization methods
for GPS samples, several security mechanisms for infrastructure and
communication channels, moreover we wrote a concise and comprehensive
privacy policy that informs the citizen about all processing steps.

The deliverable is accompanied with source code of the components in
Java, Python and JavaScript, documentation, and pre-compiled packages
of the executable parts.

\end{LGExecutiveSummary}

% add a \LGaddabbreviation{ABBR}{Explanation} for each abbreviation
\begin{LGAbbreviations}
%\LGaddabbreviation{LG}{Live+Gov}

\LGaddabbreviation{\textbf{AIDL}}
{Android Interface Description Language}

\LGaddabbreviation{\textbf{AJAX}}
{Asynchronous JavaScript and XML}

\LGaddabbreviation{\textbf{ALC}}
{Attribute Language with Complement}

\LGaddabbreviation{\textbf{API}}
{Application Programming Interface}

\LGaddabbreviation{\textbf{GeSA}}
{Geographical Semantic Analysis}

\LGaddabbreviation{\textbf{GPS}}
{Global Positioning System}

\LGaddabbreviation{\textbf{GSM}}
{Global System for Mobile Communications}

\LGaddabbreviation{\textbf{HTML}}
{HyperText Markup Language}

\LGaddabbreviation{\textbf{HTTP}}
{Hypertext Transfer Protocol}

\LGaddabbreviation{\textbf{ID}}
{Identifier}

\LGaddabbreviation{\textbf{IEEE 802.11}}
{see WIFI, WLAN}

\LGaddabbreviation{\textbf{JSON}}
{JavaScript Object Notation}

\LGaddabbreviation{\textbf{LAN}}
{Local Area Network}

\LGaddabbreviation{\textbf{REST}}
{Representational State Transfer}

\LGaddabbreviation{\textbf{RF-ID}}
{Radio-Frequency Identification}

\LGaddabbreviation{\textbf{SDCF}}
{Sensor Data Collection Framework}

\LGaddabbreviation{\textbf{SQL}}
{Structured Query Language}

\LGaddabbreviation{\textbf{SVM}}
{Support Vector Machine}

\LGaddabbreviation{\textbf{TBox}}
{Terminological Box}

\LGaddabbreviation{\textbf{UI}}
{User Interface}

\LGaddabbreviation{\textbf{URL}}
{Uniform Resource Locator}

\LGaddabbreviation{\textbf{UUID}}
{Universal Unique Device Identifier}

\LGaddabbreviation{\textbf{WP}}
{Work Package}

\LGaddabbreviation{\textbf{WIFI}}
{Wireless Fidelity (IEEE 802.11), WLAN}

\LGaddabbreviation{\textbf{WLAN}}
{Wireless Local Area Network}

\LGaddabbreviation{\textbf{XML}}
{Extensible Markup Language}

\LGaddabbreviation{~\\}
{~~~}

\end{LGAbbreviations}

% add a \LGaddterm{Term}{Explanation} for each entry of the glossary
%\begin{LGGlossary}
%\LGaddterm{Term}{Definition text here}
%\LGaddterm{Term2}{Definition text here}
%\end{LGGlossary}

\setcounter{tocdepth}{2}

% add this for a table of contents
\LGTOC

\newpage
\newcommand{\todo}[2]{\textcolor{magenta}{#1: #2}}
%\newcommand{\todo}[2]{}

\chapter{Introduction}\label{chap:Introduction}
In recent days the importance of privacy protection has been amplified
by the reports about the mass surveillance of ordinary citizens on a
global scale by the NSA and other intelligence agencies around the
world.

While aiming at the noble cause of enhancing eParticipation using
mobile technologies, Live+Gov systems do process a large variety data
that is potentially infringing the citizens privacy. The captured data
includes personal information like name, phone numbers and email
addresses and sensor data from GPS and accelerometer sensors. Also
with some applications it is possible to gather images and textual
input from the citizen.

While the collection of this data is necessary for providing the
advanced services that Live+Gov aims to deliver, at he same time, the
available raw data can be used to draw a very detailed picture of the
private life of the citizen. For instance can GPS location tracking be
used to reveal shopping habits (e.g. when a car seller is visited) and
associations to political groups (when a meeting is
attended). Accelerometer data can be used to infer medical conditions
like walking disabilities. Images can contain faces of nearby persons
to with whom the citizen is associated. All this data is highly
sensitive to the citizens privacy and can be used against the citizen
if it falls in the wrong hands.

The great importance of protecting the citizens privacy should be
apparent from these examples. The European Union, as well as many
other countries in the past, has set out a number of directives that
regulate the collection, processing and use of privacy sensitive
data. We explain the most relevant legislation in Section
\ref{sec:legal}.

The ethical aspects of privacy have been the subject of study of many
social scientists and philosophers. One scholar which is particularly
relevant in our context is Charles Fried. He investiages, why we are
intuitively so sensitive to violations of our privacy. For him privacy
is not asserted as an intrinsic value by itself, he rather sated:

\begin{quote}
Privacy is not simply an absence of information about us in the minds of others;
rather it is the control we have over information about ourselves.
\end{quote}

Fried`s study on the understanding of privacy provided a great
contribution to the research on the same term in philosophy and
computer science and despite the fact his text was published in 1970,
he already included technologies to its viewpoint (like location
monitoring) that are particularly relevant to our Context.  We explain
his theory in \ref{sec:ethics} and follow his definition of privacy in
this document.

In this document we perform a thorough analysis of how to protect the
privacy of a citizen and present several implementations that form the
core of our privacy aware Sensor Data Storage and Mining Infrastructure.

We identify six main threats for the citizens privacy, and derive
eight recommendations that should be followed in order to reduce the
associated risks for hazards.  This analysis form the guideline for
the selection of 9 measures that were implemented in our system and
documented in Chapter \ref{chap:impl}.

\input{PrivacyProtection-2014-09-09}

\clearpage

\input{Implementations}

\bibliography{D1-3}
\end{document}