### 2014-12-18 figures/LG-Implicit-Sensor-Privacy-Matrix.tex

\begin{figure}
\centering
\includegraphics[width=\textwidth]{diagrams/png/implicit-sensor-privacy-matrix.png}

%\begin{flushleft}
%\scriptsize
%\textbf{Legend:}
%\begin{itemize}
%\itemsep1pt\parskip0pt\parsep0pt
%\item
%  The x-axis lists the 7 Types of Privacy according to Friedewald et al.
%\item
%  The y-axis lists all mobile sensors currently used by the Live+Gov
%  procject.
%\item
%  The big black bullet point dentotes that there is a \textbf{direct}
%  violoation of a privacy type from a sensor
%\end{itemize}
%\end{flushleft}

\caption{Live+Gov Implicit Sensor-Privacy Matrix}
\label{figure:Live+Gov Implicit Sensor-Privacy Matrix}
\end{figure}