### 2014-12-18 llncs/paper.tex
added basic structure to paper.tex

\documentclass[runningheads,a4paper]{llncs}

\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{listings, color}
\usepackage[bookmarks,bookmarksopen,bookmarksdepth=2]{hyperref}

\newcommand{\TODO}[1]{\textcolor{magenta}{TODO: #1}}

\graphicspath{ {./img/} }

\begin{document}

%============================================================

\title{Privacy Analysis of a Mobile Sensor Application}
\author{Heinrich Hartmann \and Tim Wambach \and Maximilian Meffert \and R\"udiger Grimm}
\institute{University of Koblenz-Landau}
\maketitle

%============================================================

\begin{abstract}

\TODO{later}

\keywords{
privacy protection, IT security analysis, sensor data, mobile phones, traffic survey
}

\end{abstract}

%============================================================

\section{Introduction}

\TODO{RG}

%============================================================

\section{Scenario Description: Mobile Traffic Survey}

\TODO{HH}

%============================================================

\section{Definition of Privacy}

\TODO{MM}

%============================================================

\section{Privacy Analysis}

\TODO{TW}

%============================================================

\section{Conclusion}

\thebibliography

\end{document}